Empirical Bayes analysis of \RNAseq{} data for detection of gene expression heterosis

### 2015-08-26 ebayes_rnaseq_heterosis.tex
fixed up real data results

We used our method to analyze a maize data set \citep{paschold2012complementation} of \RNAseq{} gene expression in parental lines B73 and Mo17 and the hybrid genotype (B73$\times$Mo17). The data are available in the Supplemental Materials. Each variety had four biological replicates measured with Illumina methodology and equipment. Reads were mapped to the whole reference genome using the short reads aligner, NOVOALIGN. For more specifics, please see \cite{paschold2012complementation}.

As explained in Section~\ref{s:sim_data}, we analyzed a set of $G=27,888$ genes for $V=3$ varieties and $n_v=4$ replicates ($v=1,2,3$). We estimated the hyperparameters to be $\hat{\eta}_\phi = 3.44$, $\hat{\sigma}_\phi = 2.74$, $\hat{\eta}_\alpha = -0.05$, $\hat{\sigma}_\alpha=0.36$, $\hat{\eta}_\delta = 0.08$, $\hat{\sigma}_\delta=0.33$, $\hat{\eta}_\psi = -2.35$, and $\hat{\sigma}_\psi=0.60$.

@@

Figure \ref{f:gene_specific_estimates} provides point estimates from the initial \edgeR{} estimation step and after the eBayes procedure described in Section \ref{s:ebayes}. 
\begin{figure}
\centerline{\includegraphics[width=\textwidth]{gene_specific_estimates.pdf}}
\caption{Two-dimensional histogram of point estimates from \edgeR{} and posterior means from eBayes using Laplace distributions along with the $y=x$ diagonal.}
\label{f:gene_specific_estimates}
\end{figure}
This figure shows shrinkage for large absolute values of $\alpha$ and $\delta$ from the \edgeR{} estimation toward $\hat{\eta}_\alpha$ and $\hat{\eta}_\delta$ which are both approximately zero. The figure also shows decreased values for the overdispersion parameter with larger decreases for high and low values of overdispersion. Finally, very little, if any, shrinkage is observed for $\phi$.

>>>>

We used our method to analyze a maize data set \citep{paschold2012complementation} of \RNAseq{} gene expression in parental lines B73 and Mo17 and the hybrid genotype (B73$\times$Mo17). Each variety had four biological replicates measured with Illumina methodology and equipment. Reads were mapped to the whole reference genome using the short reads aligner, NOVOALIGN. For more specifics, please see \cite{paschold2012complementation}.

@@

For the eBayes approach with the Laplace distributions, we estimated the hyperparameters to be $\hat{\eta}_\phi = 3.44$, $\hat{\sigma}_\phi = 2.74$, $\hat{\eta}_\alpha = -0.05$, $\hat{\sigma}_\alpha=0.36$, $\hat{\eta}_\delta = 0.08$, $\hat{\sigma}_\delta=0.33$, $\hat{\eta}_\psi = -2.35$, and $\hat{\sigma}_\psi=0.60$. We then performed independent MCMC analysis for each gene conditional on these hyperparameters. As with the simulation study, we ran 4 chains simultaneously and doubled the number of MCMC iterations until each gene-specific parameter had an effective sample size above 1,000.

Figure \ref{f:gene_specific_estimates} provides point estimates of the gene-specific parameters from the initial \edgeR{} estimation step and after the eBayes with Laplace distribution procedure described in Section \ref{s:ebayes}. 
\begin{figure}
\centerline{\includegraphics[width=\textwidth]{gene_specific_estimates.pdf}}
\caption{Two-dimensional histogram of point estimates from \edgeR{} and posterior means from eBayes using Laplace distributions along with the $y=x$ diagonal.}
\label{f:gene_specific_estimates}
\end{figure}
This figure shows shrinkage for large absolute values of $\alpha$ and $\delta$ from the \edgeR{} estimation toward $\hat{\eta}_\alpha$ and $\hat{\eta}_\delta$ which are both approximately zero. The figure also shows decreased values for the overdispersion parameter with larger decreases for high and low values of overdispersion. Finally, very little, if any, shrinkage is observed for $\phi$.
